---
title: "Dice Rolls"
author: "Kit Applegate"
date: "4/4/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of Daily Dice Rolls Over Two Year Period
Rolls by Eric Prufer

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(tidymodels)
library(lubridate)
library(scales)

rolls <- read.csv("daily.Rolls.csv")

rolls$Date <- as_date(dmy(rolls$Date))

rolls$Weekdays <- weekdays(rolls$Date)

rolls$Month.Name <- months(as.Date(rolls$Date))

rolls <- rolls %>%
  dplyr::mutate(year = lubridate::year(Date), 
                month = lubridate::month(Date), 
                day = lubridate::day(Date))

```

## Distribution

First, the distribution was examined. As observed, it is pretty even across all days of the week. 

```{r Data Exploration, echo=FALSE, message=FALSE, warning=FALSE}
library(viridis)



rolls %>%
  mutate(Weekdays = factor(Weekdays, 
                           levels = c("Monday", "Tuesday", "Wednesday",
                                      "Thursday", "Friday", "Saturday",
                                      "Sunday"))) %>%
  ggplot(aes(x = Weekdays, y = Roll, fill = Weekdays)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter() +
  labs(title = "Box and Whisker Plot with a Point Plot Depicting Rolls by Weekday",
       subtitle = "Rolls by Eric Prufer",
       x = "Side Landed on When Rolled", 
       y = "Day of the Week") +
  theme_classic() +
  theme(legend.position="none")
  




```

There is more variation across months as observed in the plot below.  


```{r echo=FALSE, message=FALSE, warning=FALSE}
rolls %>%
  mutate(Month.Name = factor(Month.Name, 
                           levels = c("January","February", "March", "April", "May",
                                      "June", "July", "August", "September", "October",
                                      "November", "December"))) %>%
  ggplot(aes(x = Month.Name, y = Roll, fill = Month.Name)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter() +
  coord_flip() +
  labs(title = "Box and Whisker Plot with a Point Plot Depicting Rolls by Month",
       subtitle = "Rolls by Eric Prufer",
       x = "", 
       y = "Side Landed on When Rolled") +
  theme(legend.position="none") +
  theme_classic() +
  theme(legend.position="none")
```


Here we observe the number of times a side was rolled. On a d20 the values of each number are weighted the same. Unlike, for example, a 2d6 is inherently higher weighted to roll certain sides. We can observe there is not a pattern forming across each of the sides.     

```{r echo=FALSE, message=FALSE, warning=FALSE}

rolls %>%
  ggplot(aes(x = Roll)) +
  geom_histogram(binwidth=1, color = "#481567ff", fill = "#3cbb75ff", alpha=0.7) +
  labs(title = "Histogram of the Distribution of Rolls",
     subtitle = "Rolls by Eric Prufer", 
     x = "Side Landed on When Rolled", 
     y = "Number of Rolls") +
  theme_classic()
  
 

```

## Linear Trend

Last,  we can observe no linear trend. Meaning it is observed as random. There have been studies of the actual randomness of physical dice rolls. One study mentioned in *Inside Science* "Dice Rolls are Not Completely Random", indicates surface, which number is face up on the dice and a number of other factors can influence the outcome of the roll. For the sake of the non-scientific nature and the difficultly the regression model had predicting the outcome of the rolls, the conclusion is that it is random (enough) for us to say it is random.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(rolls, aes(x = Roll, y = day)) +
  geom_jitter() +
  geom_smooth(method=lm , color="#481567ff", fill="#3cbb75ff", se=TRUE) +
  labs(title = "Linear Trend Plot",
       subtitle = "Rolls by Eric Prufer", 
       x = "Side Landed on When Rolled", 
       y = "Day of the Month") +
  theme_classic()


```


